---
title: "Penguin ANOVA"
author: "Danielle Poole"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages: 
```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <-read_csv("penguins.csv")
```

is there a sign different in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago?

single factor Species
levels 3 
variable flipper length (not mean flipper length)

Maybe I can use a one way ANOVA
-Independent observations 
-Equal Varience
-Normality (histograms/qqplots)

```{r}
counts <-with(penguins, table(species, sex))

penguins2<-penguins%>%
  mutate(
    common_name =
      case_when(
        species =="Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>%
  select (-species) %>%
  filter(sex =="FEMALE")

```

Exploratory grapghs and Levene's test for equal variances

```{r}
hist <-ggplot(penguins2, aes(x =flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~ common_name, scale = "free")

hist

qq<-ggplot(penguins2, aes(sample = flipper_length))+
  geom_qq(aes(color=common_name))+
  facet_wrap(~common_name, scales = "free")

qq


##Based on qq-plots and histograms, observations for each group appear normally distributed. Also, n>30 for all groups, so means will be mornally distributed by CLT.
```

Evaluate variances and perform a Levenes Test for equal variances
```{r}
#If the largest sample varience is <4x greater than the smallest sample variance, then these are "close enough" to consider variences equal


variances <- penguins2 %>%
  group_by((common_name))%>%
  summarize(
    variance =var(flipper_length)
  )

variances

#The largest variance (33.1) is definitely not more than 4x greater than the smallest varience (15.2)

#Levenes Test

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin


##H0: No difference in variances (variances are equal)
#HA: Variences are NOT equal

#There is no sign difference in variences across species. 
```
One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}
penguin_aov <- aov(flipper_length ~ common_name, data= penguins2)
summary(penguin_aov)



#Post-hoc testing by Tukeys HSD
#H0: No sign diff in pairwise means
#HA: There IS a sign difference in pairwise means


penguin_ph <-TukeyHSD(penguin_aov)
penguin_ph
```


Mean flipper lengths og female penguins differed sign for all species studied (one way anova, F(df)= statistic, p<.001, $\alpha$ = 0.05 with post-hoc Tukey'sHSD, $\aphpa$ = 0.05)























